A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol

### 2023-04-08 paper.Rmd
re-working section 2.2 and 5.1, 2.2 still needs more

Other than checking diagnostic plots, analysts may perform formal hypothesis testing to detect model defects. A variety of tests can be applied; each testing for a specific violation of the null hypothesis. For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution. To test specific forms of non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the predictors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969]. The Shapiro-Wilk test [@shapiro1965analysis] is the most widely used test of non-normality included by many of the statistical software programs. The Jarque--Bera test [@jarque1980efficient] is also used to directly check whether the sample skewness and kurtosis match a normal distribution.

We apply the RESET test, Breusch-Pagan test and Shapiro-Wilk test to the residual plots shown in Figure \ref{fig:residual-plot-common-departures}; the results of this is shown in Table \ref{tab:example-residual-plot-table}. The Breusch-Pagan test and the Shapiro-Wilk test both reject $H_0$ for residual plots that display non-linearity and heteroscedasticity, even though these are not the original intention of the test. As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are also sensitive to other types of departures. It is likely $H_0$ is correctly rejected but for the wrong reason, a phenomenon known as the "Type III error". Additionally, outliers will often incorrectly trigger the rejection of $H_0$ despite when majority of the residuals are well-behaved [@cook_applied_1999]. Furthermore, with a sufficiently large sample size, residual-based tests may reject $H_0$ due to a slight departure that is of little practical significance. These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures.

@@

As discussed in Section \ref{conventionally-testing-for-departures}, many conventional tests are available for detecting residual departures. Implementation-wise, the built-in R package `stats` provides some commonly used residual-based tests, such as Shapiro-Wilk test. A more comprehensive collection of regression diagnostics tests can be found in the R package `lmtest` [@lmtest]. In terms of heteroskedasticity diagnostics, the R package `skedastic` [@skedastic] collects and implements 25 existing conventional tests published since 1961.

We pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`. Among them, RESET test is the only exact and appropriate test in this scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are approximate and inappropriate tests. Their estimated power is shown in Figure \ref{fig:polypower}. To set up the RESET test, we include different powers of fitted values as proxies. According to @ramsey_tests_1969, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient. Thus, we follow this guideline to conduct the RESET test. For the Breusch-Pagan test, the choice of predictors in the auxiliary regression is left to the user [@breusch_simple_1979]. But as @waldman1983note suggested, it is a good choice for the set of auxiliary predictors in the Breusch-Pagan test be the same as the White test. Thus, we include both $x$ and $x^2$ in the auxiliary regression.

>>>>

Many different hypothesis tests are available to detect specific model defects. For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution. To test specific forms of non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the predictors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969]. The Shapiro-Wilk test [@shapiro1965analysis] is the most widely used test of non-normality included by many of the statistical software programs. The Jarque--Bera test [@jarque1980efficient] is also used to directly check whether the sample skewness and kurtosis match a normal distribution.

Table \ref{tab:example-residual-plot-table} shows the RESET, Breusch-Pagan and Shapiro-Wilk test results ($p$-values) for the residual plots shown in Figure \ref{fig:residual-plot-common-departures}. The RESET test  and Breusch-Pagan test were computed using the `resettest` and `bptest` functions from R package `lmtest`, respectively. The Shapiro-Wilk test  was computed using the `shapiro.test` from the core R package `stats`. (The the R package `skedastic` [@skedastic] contains a large collection of tests for heteroskedasticity.) The RESET test is exact, but the power parameter must be chosen. @ramsey_tests_1969 suggests using a power of four, which is what we used. The Breusch-Pagan and Shapiro-Wilk tests are approximate.

We would expect the RESET test for non-linearity to reject residual plot B, the Breusch-Pagan test for heteroskedasticity to reject the residual plot C, and Shapiro-Wilk test for non-normality to reject residual plot D, which they all do and correctly fail to reject residual plot A. Interestingly, the Breusch-Pagan and Shapiro-Wilk tests also reject the residual plots exhibiting structure that they weren't designed for. @cook1982residuals discusses that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures, that is, correctly rejected but for the wrong reason, a phenomenon known as the "Type III error". Also, some types of departure can also have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity. Additionally, other problesms such as outliers can trigger the rejection [@cook_applied_1999]. With a sufficiently large sample size, residual-based tests may reject when there is only a slight departure. This is a situation that is addressed by the results in this paper because from a practical perspective we would only want to flag the model as problematic when ther is a XXX diagnostic  practical significance. <!--These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures. -->
 XXX check heteroscedastic spelling

<!--
As discussed in Section \ref{conventionally-testing-for-departures}, many conventional tests are available for detecting residual departures. Implementation-wise, the built-in R package `stats` provides some commonly used residual-based tests, such as Shapiro-Wilk test. A more comprehensive collection of regression diagnostics tests can be found in the R package `lmtest` [@lmtest]. In terms of heteroskedasticity diagnostics, the R package `skedastic` [@skedastic] collects and implements 25 existing conventional tests published since 1961.

We pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`. Among them, RESET test is the only exact and appropriate test in this scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are approximate and inappropriate tests. Their estimated power is shown in Figure \ref{fig:polypower}. To set up the RESET test, we include different powers of fitted values as proxies. According to @ramsey_tests_1969, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient. Thus, we follow this guideline to conduct the RESET test. For the Breusch-Pagan test, the choice of predictors in the auxiliary regression is left to the user [@breusch_simple_1979]. But as @waldman1983note suggested, it is a good choice for the set of auxiliary predictors in the Breusch-Pagan test be the same as the White test. Thus, we include both $x$ and $x^2$ in the auxiliary regression.
-->

@@

### 2023-04-08 paper.tex
re-working section 2.2 and 5.1, 2.2 still needs more

Other than checking diagnostic plots, analysts may perform formal
hypothesis testing to detect model defects. A variety of tests can be
applied; each testing for a specific violation of the null hypothesis.
For example, the presence of heteroskedasticity can usually be tested by
applying the White test \citep{white_heteroskedasticity-consistent_1980}
or the Breusch-Pagan test \citep{breusch_simple_1979}, which are both
derived from the Lagrange multiplier test \citep{silvey1959lagrangian}
principle that relies on the asymptotic properties of the null
distribution. To test specific forms of non-linearity, one may apply the
F-test as a model structural test to examine the significance of
specific polynomial and non-linear forms of the predictors, or the
significance of proxy variables as in the Ramsey Regression Equation
Specification Error Test (RESET) \citep{ramsey_tests_1969}. The
Shapiro-Wilk test \citep{shapiro1965analysis} is the most widely used
test of non-normality included by many of the statistical software
programs. The Jarque--Bera test \citep{jarque1980efficient} is also used
to directly check whether the sample skewness and kurtosis match a
normal distribution.

We apply the RESET test, Breusch-Pagan test and Shapiro-Wilk test to the
residual plots shown in Figure
\ref{fig:residual-plot-common-departures}; the results of this is shown
in Table \ref{tab:example-residual-plot-table}. The Breusch-Pagan test
and the Shapiro-Wilk test both reject \(H_0\) for residual plots that
display non-linearity and heteroscedasticity, even though these are not
the original intention of the test. As discussed in
\citet{cook1982residuals}, most residual-based tests for a particular
type of departure from model assumptions are also sensitive to other
types of departures. It is likely \(H_0\) is correctly rejected but for
the wrong reason, a phenomenon known as the ``Type III error''.
Additionally, outliers will often incorrectly trigger the rejection of
\(H_0\) despite when majority of the residuals are well-behaved
\citep{cook_applied_1999}. Furthermore, with a sufficiently large sample
size, residual-based tests may reject \(H_0\) due to a slight departure
that is of little practical significance. These can be largely avoided
in diagnostic plots as experienced analysts can evaluate the
acceptability of assumptions flexibly, even in the presence of outliers
and slight departures.

@@

As discussed in Section \ref{conventionally-testing-for-departures},
many conventional tests are available for detecting residual departures.
Implementation-wise, the built-in R package \texttt{stats} provides some
commonly used residual-based tests, such as Shapiro-Wilk test. A more
comprehensive collection of regression diagnostics tests can be found in
the R package \texttt{lmtest} \citep{lmtest}. In terms of
heteroskedasticity diagnostics, the R package \texttt{skedastic}
\citep{skedastic} collects and implements 25 existing conventional tests
published since 1961.

We pick RESET test (\texttt{resettest}) and Breusch-Pagan test
(\texttt{bptest}) from the R package \texttt{lmtest}, and Shapiro-Wilk
test (\texttt{shapiro.test}) from the built-in R package \texttt{stats}.
Among them, RESET test is the only exact and appropriate test in this
scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are
approximate and inappropriate tests. Their estimated power is shown in
Figure \ref{fig:polypower}. To set up the RESET test, we include
different powers of fitted values as proxies. According to
\citet{ramsey_tests_1969}, there are no general rules for the power of
the fitted values needed by the RESET test, but it finds power up to
four is usually sufficient. Thus, we follow this guideline to conduct
the RESET test. For the Breusch-Pagan test, the choice of predictors in
the auxiliary regression is left to the user
\citep{breusch_simple_1979}. But as \citet{waldman1983note} suggested,
it is a good choice for the set of auxiliary predictors in the
Breusch-Pagan test be the same as the White test. Thus, we include both
\(x\) and \(x^2\) in the auxiliary regression.

>>>>

Many different hypothesis tests are available to detect specific model
defects. For example, the presence of heteroskedasticity can usually be
tested by applying the White test
\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
test \citep{breusch_simple_1979}, which are both derived from the
Lagrange multiplier test \citep{silvey1959lagrangian} principle that
relies on the asymptotic properties of the null distribution. To test
specific forms of non-linearity, one may apply the F-test as a model
structural test to examine the significance of specific polynomial and
non-linear forms of the predictors, or the significance of proxy
variables as in the Ramsey Regression Equation Specification Error Test
(RESET) \citep{ramsey_tests_1969}. The Shapiro-Wilk test
\citep{shapiro1965analysis} is the most widely used test of
non-normality included by many of the statistical software programs. The
Jarque--Bera test \citep{jarque1980efficient} is also used to directly
check whether the sample skewness and kurtosis match a normal
distribution.

Table \ref{tab:example-residual-plot-table} shows the RESET,
Breusch-Pagan and Shapiro-Wilk test results for the residual plots shown
in Figure \ref{fig:residual-plot-common-departures}. The RESET test and
Breusch-Pagan test were computed using the \texttt{resettest} and
\texttt{bptest} functions from R package \texttt{lmtest}, respectively.
The Shapiro-Wilk test was computed using the \texttt{shapiro.test} from
the core R package \texttt{stats}. (The the R package \texttt{skedastic}
\citep{skedastic} contains a large collection of tests for
heteroskedasticity.) The RESET test is exact, but the power parameter
must be chosen. \citet{ramsey_tests_1969} suggests using a power of
four, which is what we used. The Breusch-Pagan and Shapiro-Wilk tests
are approximate.

The Breusch-Pagan test and the Shapiro-Wilk test both reject \(H_0\) for
residual plots that display non-linearity and heteroscedasticity, even
though these are not the original intention of the test. As discussed in
\citet{cook1982residuals}, most residual-based tests for a particular
type of departure from model assumptions are also sensitive to other
types of departures. It is likely \(H_0\) is correctly rejected but for
the wrong reason, a phenomenon known as the ``Type III error''.
Additionally, outliers will often incorrectly trigger the rejection of
\(H_0\) despite when majority of the residuals are well-behaved
\citep{cook_applied_1999}. Furthermore, with a sufficiently large sample
size, residual-based tests may reject \(H_0\) due to a slight departure
that is of little practical significance. These can be largely avoided
in diagnostic plots as experienced analysts can evaluate the
acceptability of assumptions flexibly, even in the presence of outliers
and slight departures.

@@
